Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-18 paper.tex
improve puinv chart a bit
a bit more discussion about drawdown results

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps and individual reactor modeling
create a need for larger floating inventory buffers in order to avoid material
shortages.  Although direct measures of this effect are difficult, some of its
consequences can be quantified by comparing scenarios with different time step
durations (e.g. cases 2 and 4).

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are
primarily caused by the \emph{inventory drawdown effect}.  Figure
\ref{fig:puinv} helps to visualize this effect.  The black dots in
the figure represent impulse flows of available Pu out of the fuel fab and
into fast reactors.  Blue lines represent the flows into Pu inventory
available for making fuel. The in flows are not shown for the individual
reactor cases because they are somewhat messy and obscure otherinformation in
the plots. The colored curves show Pu inventory available for making fast
reactor fuel.  When the inventory curves meet and force down the flows, fuel
shortages occur.

Quarterly time steps have less frequent but larger impulse material flows.
These larger flows, however, do not compensate for the larger withdrawals
because facilities do not know about incoming inventory when they resolve
their outflow for a particular time step.  This information lag effectively
requires the floating Pu inventory to be maintained at a higher level.  This
can be clearly seen in the case 4 plot which has a higher standing inventory
than case 2 during the fuel shortage years.  The case 4 Pu inventory is unable
to drop below approximately 100 tonnes, which indicates out-flows to be
roughly equal to the Pu flow into inventory each time step during those years.
This can be confirmed by the in-flow and inventory curves touching at those
points.  Case 2 Pu inventory is similarly limited at about the 30 tonne level.
As expected, the case 4 withdrawals are also three times higher case 2
withdrawals.

>>>>

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
correspondingly larger inventory top-up quantities.  The other component of
this effect is the simultaneity of incoming and outgoing material flows on a
given time step.  Within a particular time step facilities do not know about
incoming material when they make commitments for outgoing material.  For
longer time steps, larger incoming quanta of material are not available for
making offers. In general longer time steps and individual reactor modeling
create a need for larger floating inventory buffers in order to avoid material
shortages.  Although direct measures of this effect are difficult, some of its
consequences can be quantified by comparing scenarios with different time step
durations (e.g. cases 2 and 4) such as higher standing inventories during
shortages.

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are
primarily caused by the \emph{inventory drawdown effect}.  Figure
\ref{fig:puinv} helps to visualize this effect.  The black dots in
the figure represent impulse flows of available Pu out of the fuel fab and
into fast reactors.  Blue lines represent the flows into Pu inventory
available for making fuel. The in flows are not shown for the individual
reactor cases because they are somewhat messy and obscure other information in
the plots. The colored curves show Pu inventory available for making fast
reactor fuel.  When the inventory curves meet and force down the flows, fuel
shortages occur.

Quarterly time steps have less frequent but larger impulse material flows.
These larger flows, however, do not compensate for the larger withdrawals
because facilities do not know about incoming inventory when they resolve
their outflow for a particular time step.  This information lag effectively
requires the floating Pu inventory to be maintained at a higher level during
the shortage.  This can be clearly seen in the case 4 plot which has a higher
standing inventory than case 2 during the fuel shortage years.  The case 4 Pu
inventory is unable to drop below approximately 100 tonnes, which indicates
out-flows to be roughly equal to the Pu flow into inventory each time step
during those years.  This can be confirmed by the in-flow and inventory curves
touching at those points.  Case 2 Pu inventory is similarly limited at about
the 30 tonne level.  As expected, the case 4 withdrawals are also three times
higher case 2 withdrawals.

Throughout most of the shortage, the aggregate outflow of Pu from separated
inventory is roughly the same in cases 2 and 4.  However, due to the longer
time steps in case 4, the shortage starts slightly sooner because Pu inventory
must be maintained at a higher limit as described above.  The small difference
in outage start time has feed-back effects.  Not only are extra reactors
offline, but those extra reactors are also discharging less aggregate spent
fuel for recycling.  This farther increases the discrepency between the two
cases. Eventually the shortage in case 2 ends slightly before it does in case
4. This effect feedback occurs because the reactors in question have a
conversion ratio greater than 1.0.
